---
title: 'Coursera_JHU_Developing-Data-Products_Week2'
author: "Anik"
date: ""
output:
    html_document:
        keep_md: true
---
```{r,echo=FALSE,warning=FALSE,message=FALSE}
# load libraries
library(dplyr)
library(lubridate)
library(stringr)
library(reshape2)

library(leaflet)

```

# DATE : `r Sys.Date()`  

```{r,cache=TRUE,echo=FALSE}
#work with data

## read the data
df <- read.csv("repdata_data_StormData.csv.bz2",stringsAsFactors = FALSE)
df <- tbl_df(df)

## selecting the data
df_new <- df %>% 
    select(REFNUM, EVTYPE,BGN_DATE, FATALITIES, INJURIES, STATE) %>% 
    mutate(BGN_DATE = mdy_hms(BGN_DATE)) %>% 
    filter(year(BGN_DATE) >= 1996)

## filtering non-killers :p
logical <- (df_new$FATALITIES == 0 & df_new$INJURIES == 0)
df_new <- df_new %>% filter(!logical)

## keeping backup
backup_df_new <- df_new

# cleaning data
df_new$EVTYPE = toupper(df_new$EVTYPE)
df_new$EVTYPE = str_trim(df_new$EVTYPE)

df_new[grep("TORNADO|LANDSPOUT",
            df_new$EVTYPE),]$EVTYPE <- toupper("Tornado")

## I am only interested in Tornado so lets filter it out
df_new <- df_new %>% filter(EVTYPE == toupper("Tornado"))

## reading state details
df_state <- read.csv("statelatlong.csv")

names(df_state) <- c("STATE",names(df_state)[-1])

df_plot <- merge(df_new,df_state,by = "STATE")

## lets selct the needed data for ploting
df_plot <- df_plot %>% 
    select(City,FATALITIES,INJURIES,Latitude,Longitude) %>% 
    group_by(City) %>% 
    summarise(TOTAL_FATALITIES = sum(FATALITIES),
              TOTAL_INJURIES = sum(INJURIES),
              LONG = mean(Longitude),
              LAT = mean(Latitude)) %>% 
    mutate(content = paste(sep = " | ", City,
                           paste("Total Injuries : ",TOTAL_INJURIES),
                           paste("Total fatalities : ",TOTAL_FATALITIES)))


leaflet(df_plot) %>% 
    setView(-96, 37.8, 4) %>% 
    addTiles() %>%
    addCircles(lng = ~LONG, lat = ~LAT,label= ~content, 
               weight = 1,radius = ~TOTAL_INJURIES * 100, 
               popup = ~City,color = "Blue") %>% 
    addCircles(lng = ~LONG, lat = ~LAT, label= ~content, weight = 1,
               radius = ~TOTAL_FATALITIES * 400, popup = ~City,color = "red") %>% 
    addLegend(position = 'bottomleft',colors = c("blue","red"),
              labels = c("Total Injuries","Total fatalities"), opacity = 1,
              title = 'Legend')
```
  
The above chart shows the total no of injuries and fatalities from Tornado’s in US from 1996 to 2011. It is clear that the south east region of US is most effected by Tornado than the rest of the country. 

For more info check out my [Exploring the U.S. National Oceanic and Atmospheric Administration’s (NOAA) storm database](https://rpubs.com/Anik_Mallick/EDA_of_NOAA).  

Data Source: [Storm Data [47Mb]](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)  

[USA lat,long for state abbreviations](https://www.kaggle.com/washimahmed/usa-latlong-for-state-abbreviations/downloads/usa-latlong-for-state-abbreviations.zip/1)

Code for this analysis can be found here : [Code](https://github.com/AnikMallick/DDP_Coursera/tree/master/Week%202)
